---
title: "2021 Bacterial Abundance"
author: "Ryan Pavlovich"
date: "10/20/2021"
output: github_document
---
# Goal 

This document shows how **individual bottle** bacterial abundance data from 2021 remineralization experiments were processed, QC'd and analyzed. It also provides an intro to data processing and analysis with Rstudio and R Markdown.

#Intro to R Markdown

```{r Load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

# Import Data

```{r}
excel_sheets("~/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx")

metadata <- read_excel("~/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx", sheet = "Metadata")

glimpse(metadata)
#unique(metadata$Bottle)
#unique(metadata$Treatment)

data <- read_excel("~/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx", sheet = "Data")
glimpse(data)

joined <- left_join(metadata, data)
glimpse(joined)

```

# Prepare Data

Date and Time column values were converted from characters to dates, while adding column with time elapsed for each treatment, as well as converting to cells/L. This will help us match up with the TOC data later. We will then subset the data for variables of interest and drop NA value. 

To do this, we are going to be using **piping**. Piping is an operation that allows to write more efficient code. The way that we'll use it here is to manipulate our data sequentially.

```{r message=FALSE, warning=FALSE}
cells <- joined %>%
  mutate(Datetime = ymd_hm(Datetime)) %>%
  mutate(all_cells_L = as.numeric(all_cells_uL) * 1000000) %>%
  group_by(Treatment, Bottle) %>%
  mutate(interv = interval(first(Datetime), Datetime), 
         s = as.numeric(interv), 
         hours = s/3600, 
         days = hours/24) %>%
  ungroup() %>%
  select(Experiment:DNA_Sample, all_cells_L, hours, days) %>%
  drop_na(all_cells_L)
glimpse(cells)
```
# Plot Growth Curves

#Growth curve plotted for each bottle

```{r}
#assign hex colors to our different treatments
custom.colors <- c("Control" = "#377EB8", "Kelp Exudate" = "#4DAF4A", "Kelp Exudate_Nitrate_Phosphate" = "#E41A1C", "Glucose_Nitrate_Phosphate" = "#FF7F00")

levels <- c("Control", "Kelp Exudate", "Kelp Exudate_Nitrate_Phosphate", "Glucose_Nitrate_Phosphate")

cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=all_cells_L, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size=1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  theme_light()
glimpse(cells)

```


# Next Steps

We can calculate:

- total change in cells from initial condition to the end of the experiment.
- specific growth rate as the slope of ln(abundance) v time during exponential growth phase
- doubling time as ln(2) divided by the specific growth rate
- mean of each of these parameters across each treatment

1st, we need to determine **where** exponential growth is occurring in each of our bottles, if it does. To do this, we'll plot ln(abundance) vs time.

# Identify exponential phase of growth in our remin experiments

```{r}
ln_cells <- cells %>%
  group_by(Treatment, Bottle) %>%
  mutate(ln_cells = log(all_cells_L),
         diff_ln_cells = ln_cells - lag(ln_cells, default =
first(ln_cells)))
```

```{r fig.height=6, fig.width=8}
ln_cells %>%
    mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=diff_ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size=1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("∆ln cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_wrap("Bottle", ncol =2) +
  theme_light()
```
Exponential growth seems to occur right at the beginning of the experiment between 0-1 days for the most of the bottles in samples with abundance data points for those days. There is also some exponential growth between 3-4 days for a majority of the bottles. Note that some of the abundance timepoints are missing due to sample labels falling off in LN2.

Let's try plotting ln_cells to see if that can help us identify exponential growth in the control a little better. 

```{r fig.height=6, fig.width=8}
ln_cells %>%
    mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size=1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("ln cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_wrap("Bottle", ncol =2) +
  theme_light()

```




